---
title: "HW04 Part 1: Complete Chapter 3"
author: "Andrew Chronister"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 3

- Change "your name" in the YAML header above to your name.

### Section 3.1

Load the tidyverse packages. I've entered the first code chunk for you. **You must enter all subsequent code chunks and run the code.** 

```{r}
library("tidyverse")
```

If you get an error, then you did not install the `tidyverse` package like you were supposed to in the previous assignment. Go back to that assignment and complete that step. (This should not be the case, though.)

### Section 3.2: First steps

Enter your code chunks for Section 3.2 here.

#checking out the mpg data
```{r}
mpg
```

#checking out the ggplot2 results, looks the same. 


```{r}
ggplot2::mpg
```

#running the ggplot code to create a scatterplot of mpg vs displacement
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy))
```

Get in the habit of entering a brief description of what each chunk of code does. Enter this above the chunk, like I did in Section 3.1.

#### Section 3.2 Questions

Answer the questions *completely.* Some answers will require you to write code in a code chunk and also to type text before or after the chunk. Other answers will require only code or only text. 

**1:** Run ggplot(data = mpg). What do you see?
```{r}
ggplot(data = mpg)
```
#All I see when I run ggplot(data=mpg) is a blank box.  That's not very interesting.  I assume that in order to see any of the data, you have to tell ggplot what kind of geometric function, mapping and aesthetic system it is supposed to use, otherwise it just shrugs its shoulders and shows exactly nothing. 

**2:** How many rows are in `mpg`? How many columns? **Hint:** Use the `dim()` function. Type `dim()` in the console to learn what `dim()` does.
#It appears that there are 234 rows of data with 11 different columns.  

**3:** What does the `drv` variable describe? Read the help for `?mpg` to find out.
```{r}
?mpg
```
#drv represents whether the vehicle is front wheel, rear wheel or 4-wheel drive.

**4:** Make a scatterplot of `hwy` vs `cyl`.
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=hwy, y=cyl))
```
#that plot makes sense.  


**5:** What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?

```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=class, y=drv))
```
#The plot is not useful because all of the results are stacked on top of each other at each intersection. In other words, the intersection of compact class and 4 drv may have multiple vehicles represented, but vecause they are stacked on top of each other, you can't tell.  


### Section 3.3: Aesthetic mappings
#mapping mpg data using color aesthetic for class. 
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, color=class))
```

#mapping mpg data using the size=class aesthetic.
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, size=class))
```

#using class to alpha aesthetic, then class to shape.
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, alpha=class))
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, shape=class))
```
#the SUV class catagory wasn't mapped because only 6 different shapes can be used at a time.  

#making all the points blue...
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy), color="blue")
```





#### Section 3.3 questions

**1:** What’s gone wrong with this code? Why are the points not blue?
#the line '"Color = "Blue"' should be outside of the aes parenthesis.  In other words, in order to make all of the plot points blue, the code should appear as:
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy), color="blue")
```
#...with a comma after the close parenthesis, and then the modifier. 

**2:** Which variables in `mpg` are categorical? Which variables are continuous? (*Hint:* type `?mpg` to read the documentation for the dataset). How can you see this information when you run mpg? 
#I would say that "class", "fl", "drv", "trans" and "model" are all categorical variables because they have no numeric value.  "displ", "year", "cyl", "cty", "hwy" all have numeric value and would be continuous variables. 


**3:** Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables? **Note:** You need several code chunks here to answer this question.
#mapping with color set to year.
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, color=year))
```
#size set to number of cylinders...
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, size=cyl))
```
#mapping with city milage set to shape. this should be interesting.  :D
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, shape=cty))
```
#well, that didn't go well.  continuous variables can't be mapped to shapes.  

#mapping color to catagorical variables:
#color = fuel
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, color=fl))
```
#now plotting size to drive type
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, size=drv))
```
#lastly, shape mapped with type of transmission
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, shape=trans))
```
#whoa, didn't realize there would be that many categories of transmission.  I'll plot it again using fuel type to shape.
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, shape=fl))
```


**4:** What happens if you map the same variable to multiple aesthetics?


**5:** What does the stroke aesthetic do? What shapes does it work with? (*Hint:* use `?geom_point`)



**6:** What happens if you map an aesthetic to something other than a variable name, like `aes(colour = displ < 5)`?


### Section 3.5: Facets



#### Section 3.5 questions

**1:** What happens if you facet on a continuous variable? **Note:** Write a prediction, then make and run your code chunk.



**2:** What do the empty cells in plot with `facet_grid(drv ~ cyl)` mean? How do they relate to this plot?



**3:** What plots does the following code make? What does `.` do?



**4:** Take the first faceted plot below. What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?



**5:** Read `?facet_wrap`. What does `nrow` do? What does `ncol` do? What other options control the layout of the individual panels? Why doesn’t `facet_grid()` have nrow and ncol argument?


**6:** When using `facet_grid()` you should usually put the variable with more unique levels in the columns. Why?

### Section 3.6: Geometric objects



#### Section 3.6.1 Questions

**1:** What geom would you use to draw a line chart? A boxplot? A histogram? An area chart? **Note:** You will have to make some types of these charts in Part 2 of this assignment.


**2:** Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.



**3:** What does `show.legend = FALSE` do? What happens if you remove it?
**Note:** Skip the "Why do you think I used it earlier in the chapter?" question.


**4:** What does the `se` argument to `geom_smooth()` do?


**5:** Will these two graphs look different? Why/why not?


**6:** Recreate the R code necessary to generate the following graphs. **Note:** Enter a separate code chunk for each graph, numbered 1 to 6, starting row-wise at upper left. 

1: upper left.


2: upper right.


3: middle left.


4: middle right.


5: lower left.


6: lower right. *Hint:* Read the help file for `geom_point`. Use `shape = 21`, `stroke = 2`, and `size = 3`.

### Section 3.7: Statistical transformations.

**Skip this section.** We may return to it later.


### Section 3.8 Position adjustments


#### Section 3.8 Questions

**1:** What is the problem with this plot? How could you improve it. *Hint:* Remember the size of the `mpg` data set from earlier in this exercise. Does the number of points seem correct? **Note:** Write the code to improve it. 


**2:** What parameters to `geom_jitter()` control the amount of jittering?


**3:** Compare and contracts `geom_jitter()` with `geom_count()`.


**4:** What is the default position adjustement for `geom_boxplot()`? Create a visualization of the `mpg` dataset that demonstrates it.


### Section 3.9: Coordinate systems

**Note:** Read about and do only the code chunk associated with `coord_flip()`. `coord_flip()` is a great layer to use when you have categorical data with longish names as it makes the text easier to read.

### Section 3.10: The layered grammar of graphics

Just read this section as a contextual review of everything you just learned.


