---
title: "HW04 Part 1: Complete Chapter 3"
author: "Andrew Chronister"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 3

- Change "your name" in the YAML header above to your name.

### Section 3.1

Load the tidyverse packages. I've entered the first code chunk for you. **You must enter all subsequent code chunks and run the code.** 

```{r}
library("tidyverse")
```

If you get an error, then you did not install the `tidyverse` package like you were supposed to in the previous assignment. Go back to that assignment and complete that step. (This should not be the case, though.)

### Section 3.2: First steps

Enter your code chunks for Section 3.2 here.

#checking out the mpg data
```{r}
mpg
```

#checking out the ggplot2 results, looks the same. 


```{r}
ggplot2::mpg
```

#running the ggplot code to create a scatterplot of mpg vs displacement
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy))
```

Get in the habit of entering a brief description of what each chunk of code does. Enter this above the chunk, like I did in Section 3.1.

#### Section 3.2 Questions

Answer the questions *completely.* Some answers will require you to write code in a code chunk and also to type text before or after the chunk. Other answers will require only code or only text. 

**1:** Run ggplot(data = mpg). What do you see?
```{r}
ggplot(data = mpg)
```
#All I see when I run ggplot(data=mpg) is a blank box.  That's not very interesting.  I assume that in order to see any of the data, you have to tell ggplot what kind of geometric function, mapping and aesthetic system it is supposed to use, otherwise it just shrugs its shoulders and shows exactly nothing. 

**2:** How many rows are in `mpg`? How many columns? **Hint:** Use the `dim()` function. Type `dim()` in the console to learn what `dim()` does.
#It appears that there are 234 rows of data with 11 different columns.  

**3:** What does the `drv` variable describe? Read the help for `?mpg` to find out.
```{r}
?mpg
```
#drv represents whether the vehicle is front wheel, rear wheel or 4-wheel drive.

**4:** Make a scatterplot of `hwy` vs `cyl`.
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=hwy, y=cyl))
```
#that plot makes sense.  


**5:** What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?

```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=class, y=drv))
```
#The plot is not useful because all of the results are stacked on top of each other at each intersection. In other words, the intersection of compact class and 4 drv may have multiple vehicles represented, but vecause they are stacked on top of each other, you can't tell.  


### Section 3.3: Aesthetic mappings
#mapping mpg data using color aesthetic for class. 
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, color=class))
```

#mapping mpg data using the size=class aesthetic.
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, size=class))
```

#using class to alpha aesthetic, then class to shape.
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, alpha=class))
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, shape=class))
```
#the SUV class catagory wasn't mapped because only 6 different shapes can be used at a time.  

#making all the points blue...
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy), color="blue")
```





#### Section 3.3 questions

**1:** What’s gone wrong with this code? Why are the points not blue?
#the line '"Color = "Blue"' should be outside of the aes parenthesis.  In other words, in order to make all of the plot points blue, the code should appear as:
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy), color="blue")
```
#...with a comma after the close parenthesis, and then the modifier. 

**2:** Which variables in `mpg` are categorical? Which variables are continuous? (*Hint:* type `?mpg` to read the documentation for the dataset). How can you see this information when you run mpg? 
#I would say that "class", "fl", "drv", "trans" and "model" are all categorical variables because they have no numeric value.  "displ", "year", "cyl", "cty", "hwy" all have numeric value and would be continuous variables. 


**3:** Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables? **Note:** You need several code chunks here to answer this question.
#mapping with color set to year.
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, color=year))
```
#size set to number of cylinders...
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, size=cyl))
```
#mapping with city milage set to shape. this should be interesting.  :D
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, shape=cty))
```
#well, that didn't go well.  continuous variables can't be mapped to shapes.  

#mapping color to catagorical variables:
#color = fuel
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, color=fl))
```
#now plotting size to drive type
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, size=drv))
```
#lastly, shape mapped with type of transmission
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, shape=trans))
```
#whoa, didn't realize there would be that many categories of transmission.  I'll plot it again using fuel type to shape.
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, shape=fl))
```


**4:** What happens if you map the same variable to multiple aesthetics?
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, alpha=class, color=class, size=class))
```
#It appears that all of the aesthetics will apply to the variable.  

**5:** What does the stroke aesthetic do? What shapes does it work with? (*Hint:* use `?geom_point`)
```{r}
?geom_point
```
#It appears that the stroke aesthetic will work on shapes theat have a border. by adjusting the stroke setting, you adjust the width of the border of the shape.  Shapes 21-24 all have a shape with a border. The shape can be filled with a certain color using the fill aesthetic and the border can be adusted with the stroke aesthetic. 

**6:** What happens if you map an aesthetic to something other than a variable name, like `aes(colour = displ < 5)`?
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy, color=displ <5))
```
#Applying the color=display<5 line performed a true or false calculation on the variables. Is the displacement less than 5?  If true, make the color one color, and false is another color.

### Section 3.5: Facets
#trying out facets...
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy))+
  facet_wrap(~class, nrow=2)
```
#faceting on two variables, drive and cylinders:
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy))+
  facet_grid(drv~cyl)
```



#### Section 3.5 questions

**1:** What happens if you facet on a continuous variable? **Note:** Write a prediction, then make and run your code chunk.
#My prediction is either that an error will occur, or it will create multiple plots with single points plotted on them, depending on how many discrete values there are.  We'll see.  
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy))+
  facet_wrap(~cty, nrow = 2)
```




**2:** What do the empty cells in plot with `facet_grid(drv ~ cyl)` mean? How do they relate to this plot?
#The empty cells in the drv~cyl plot mean that there are no vehicles that meet that criteria.  Another way to look at it, no vehicles meet those categories.  For instance, there are very few cars with 5 cylinder engines. The only vehicles that use 5 cylinder engines in this data set happen to be front wheel drive.  There are no rear weel drive or 4 wheel drive 5 cyl engine vehicles in any displacement, so there are no points to plot.  Same goes for 4 cyl rear wheel drive vehicles.  There are none in the dataset. 
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=drv, y=cyl))
```
#The relation to this plot is that the drv~cyl plot is based on categorical data in the drv category, and continuous data in cyl with a limited number. cyl almost acts as another categorical data set because the vehicles all fall in one of four cyl types.  



**3:** What plots does the following code make? What does `.` do?
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy))+
  facet_grid(drv~. )
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy))+
  facet_grid( .~cyl)
```
#the "." makes the facet_grid only create columns or rows based on the data in the specified variable instead of creating separate plots based on two variables.  


**4:** Take the first faceted plot below. What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```
#The benefits to using the facet over the color aesthetic is that multiple colors in close proximity may be difficult or impossible to distinguish.  by faceting, it is much easier to see the different groupings. The disadvantage is that now you are dealing with seven different plots instead of one.  With a larger dataset, the color aesthetic would be harder and harder to justify because the plot would appear as a colorful blob.  


**5:** Read `?facet_wrap`. What does `nrow` do? What does `ncol` do? What other options control the layout of the individual panels? Why doesn’t `facet_grid()` have nrow and ncol argument?
```{r}
?facet_wrap
?facet_grid
```
#nrow and ncol fix the number of rows and columns.  Other options that control the output include scales and shrink, which allow the scale to change depending on the output of the statistics.  There is also a drop feature that allows you to drop unused factor levels. 

#the Facet_grid function does not have nrow and ncol arguement because it forms a matrix of two variables and they will fall in a grid based on those variables.


**6:** When using `facet_grid()` you should usually put the variable with more unique levels in the columns. Why?
#I don't know. I'll try it and see what happens...
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy))+
  facet_grid(drv~class)

ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy))+
  facet_grid(class~drv)
```
#The vertical nature of the first plot looks and reads better than the horizontal nature of the second plot.  It also works better due to the rectangular nature of the screen that the plot is being viewed on.  There is more "real estate" across the top of the screen than there is down the side of the screen, therefore, you can view more columns than you can rows.  

### Section 3.6: Geometric objects
#The two plots are similar in that they represent the data, however, the smooth line on the right shows the mean and the variance, where the scatterplot shows the individual values plotted. 
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy))

ggplot(data = mpg)+
  geom_smooth(mapping = aes(x=displ, y=hwy))
```
#adding the linetype function...
```{r}
ggplot(data = mpg)+
  geom_smooth(mapping = aes(x=displ, y=hwy, linetype=drv))
```
#examples of using group and colors in geom_smooth
```{r}
ggplot(data = mpg)+
  geom_smooth(mapping = aes(x=displ, y=hwy))

ggplot(data = mpg)+
  geom_smooth(mapping = aes(x=displ, y=hwy, group=drv))

ggplot(data = mpg)+
  geom_smooth(mapping = aes(x=displ, y=hwy, color=drv), show.legend = FALSE
)
```
#plotting multiple geoms in same plot...
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy))+
  geom_smooth(mapping = aes(x=displ, y=hwy))
```
#global mapping settings.  
```{r}
ggplot(data = mpg, mapping = aes(x=displ, y=hwy))+
  geom_point()+
  geom_smooth()
```
#now adding some local settings as well.
```{r}
ggplot(data = mpg, mapping = aes(x=displ, y=hwy))+
  geom_point(mapping=aes(color=class))+
  geom_smooth()
```
#Filter introduction
```{r}
ggplot(data = mpg, mapping = aes(x=displ, y=hwy))+
  geom_point(mapping=aes(color=class))+
  geom_smooth(data = filter(mpg, class=="subcompact"), se=FALSE)
```


#### Section 3.6.1 Questions

**1:** What geom would you use to draw a line chart? A boxplot? A histogram? An area chart? **Note:** You will have to make some types of these charts in Part 2 of this assignment.
#Depending on the number of variables involved and whether continuous or categorical, a line chart can be made with geom_path or geom_freqpoly, a boxplot can be created with geom_boxplot, a histogram can be created with geom_histogram,and an area chart can be created with geom_area.  

**2:** Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.
#I believe it will create a scatterplot with different colors representing drive type. The smooth geom will create a line that represents the same data.  Not sure about the se yet...
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se=FALSE)
```
#ah, se=False turns off the range or variance feature of smooth.

**3:** What does `show.legend = FALSE` do? What happens if you remove it?
**Note:** Skip the "Why do you think I used it earlier in the chapter?" question.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(show.legend = FALSE)
```
#show.legend = toggles the legend on and off next to the plot.

**4:** What does the `se` argument to `geom_smooth()` do?
#toggling se on and off shows and hides the grey shaded areas that show varience from the mean or center of the line.


**5:** Will these two graphs look different? Why/why not?
#I believe the two graphs will be the same.  One applies the same information to both, while the second spells out the data, mappping and variable information in each geom function.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```


**6:** Recreate the R code necessary to generate the following graphs. **Note:** Enter a separate code chunk for each graph, numbered 1 to 6, starting row-wise at upper left. 

1: upper left.
```{r}
ggplot(data = mpg, mapping = aes(x=displ, y=hwy))+
  geom_point(size=3)+
  geom_smooth(se=FALSE)
```
#It appeared that the size of the scatterplot points in the example was set larger than the default, but didn't vary, so I just set the size to three to make it look similar.  

2: upper right.
```{r}
ggplot(data = mpg, mapping = aes(x=displ, y=hwy))+
  geom_point(size = 3)+
  geom_smooth(se=FALSE,mapping=aes(group = drv))
```
#took a little trial and error on that one to be honest.

3: middle left.
```{r}
ggplot(data = mpg, mapping = aes(x=displ, y=hwy, color=drv))+
  geom_point(size=3)+
  geom_smooth(se=FALSE)
```


4: middle right.
```{r}
ggplot(data=mpg, mapping = aes(x=displ, y=hwy))+
  geom_point(size=3, mapping=aes(color=drv))+
  geom_smooth(se=FALSE)
```


5: lower left.
```{r}
ggplot(data = mpg, mapping = aes(x=displ, y=hwy))+
  geom_point(size=3, mapping = aes(color=drv))+
  geom_smooth(se=FALSE, mapping = aes(linetype=drv))
```


6: lower right. *Hint:* Read the help file for `geom_point`. Use `shape = 21`, `stroke = 2`, and `size = 3`.
```{r}
ggplot(data = mpg, mapping = aes(x=displ, y=hwy))+
  geom_point(size=3, stroke=2, color="white", shape=21, mapping = aes(fill=drv))
  
```
#Still doesn't look exactly like the example.  The stroke appears to cover other data points.  

### Section 3.7: Statistical transformations.

**Skip this section.** We may return to it later.


### Section 3.8 Position adjustments
```{r}
ggplot(data=diamonds)+
  geom_bar(mapping = aes(x=cut, color=cut))
ggplot(data = diamonds)+
  geom_bar(mapping = aes(x=cut, fill=cut))
```
```{r}
ggplot(data = diamonds)+
  geom_bar(mapping = aes(x=cut, fill=clarity))
```
```{r}
ggplot(data = diamonds, mapping = aes(x=cut, fill=clarity))+
  geom_bar(alpha=1/5, position="identity")
```
```{r}
ggplot(data = diamonds, mapping = aes(x=cut, color=clarity))+
  geom_bar(fill=NA, position="identity")
```
```{r}
ggplot(data = diamonds)+
  geom_bar(mapping = aes(x=cut, fill=clarity), position="fill")
```
```{r}
ggplot(data = diamonds)+
  geom_bar(mapping = aes(x=cut, fill=clarity), position="dodge")
```
```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x=displ, y=hwy), position = "jitter")
```


#### Section 3.8 Questions

**1:** What is the problem with this plot? How could you improve it. *Hint:* Remember the size of the `mpg` data set from earlier in this exercise. Does the number of points seem correct? **Note:** Write the code to improve it. 
#Many points are mapped on top of each other, which makes this graphical representation of the data misleading.  By adding jitter, a more "accurate" image of the data can be found, (unless one looks closely at the data and sees that noise has been added to move the points apart slightly so they can be seen.)
```{r}
ggplot(data = mpg, mapping = aes(x=cty, y=hwy))+
  geom_jitter()
```



**2:** What parameters to `geom_jitter()` control the amount of jittering?
#by reading the ?geom_jitter help, width and height parameters control the amount of jittering. The default is 40%.  

**3:** Compare and contracts `geom_jitter()` with `geom_count()`.
```{r}
ggplot(data = mpg, mapping = aes(x=cty, y=hwy))+
  geom_count()
```
#geom_count counts the number of observations at each location then plots the count to the area, in the example above, it presents the count by altering the size aesthetic, so that the larger the size of the point, the more observations appear at that point.  geom_jitter on the other hand, simply adds noise to the x,y location of each observation in order to separate them from each other so that the plotted points can be seen.  

**4:** What is the default position adjustement for `geom_boxplot()`? Create a visualization of the `mpg` dataset that demonstrates it.
#If I understand the question correctly, the default position for boxplot seems to be "dodge2", which "preserves the vertical position of geom while adusting the horizontal position." according to the help.    
```{r}
ggplot(data = mpg, mapping =aes(x=cty, y=hwy))+
  geom_boxplot()
```


### Section 3.9: Coordinate systems

**Note:** Read about and do only the code chunk associated with `coord_flip()`. `coord_flip()` is a great layer to use when you have categorical data with longish names as it makes the text easier to read.

```{r}
ggplot(data = mpg, mapping = aes(x=class, y=hwy))+
  geom_boxplot()
ggplot(data = mpg, mapping = aes(x=class, y=hwy))+
  geom_boxplot()+
  coord_flip()
```
#skipping the quickmap, polar and exercises from this section.

### Section 3.10: The layered grammar of graphics

Just read this section as a contextual review of everything you just learned.
#Done... whew.  This assignment sure seemed long.  Maybe it was the egg nog.  :D

