---
title: "HW05 Part 1: Complete the sections"
author: "Andrew Chronister"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- Change "your name" in the YAML header above to your name.

- As usual, enter the examples in code chunks and run them, unless told otherwise.

## Chapter 10: Tibbles

Read [R4ds Chapter 10: Tibbles](https://r4ds.had.co.nz/tibbles.html), sections 1-3.

### 10.1: Introduction

Load the tidyverse package. 
```{r}
library(tidyverse)
```

### 10.2: Creating tibbles
```{r}
iris
```

Enter your code chunks for Section 10.2 here.
```{r}
as.tibble(iris)
```
#the iris data is displayed in a tibble, but it appears just as it does if I type iris in a code chunk.  I'm sure that is going to change...  
Describe what each chunk code does. 
```{r}
tibble(
  x=1:5,
  y=1, 
  z=x^2+y
)
```
#This tibble creates 3 new vectors, x which equals 1-5, y which equals 1 and z which is x squared plus y.  

```{r}
tb <- tibble(
  `:)`="smile",
  ` ` ="space",
  `2000`="number"
)
tb
```
#This example shows how the column names can be virtually anything, even a space.  Using backticks allows us to name variables.
```{r}
tribble(
  ~x, ~y, ~z,
  #--|--|---- Header comment as suggested in example
  "a", 2, 3.6,
  "b", 1, 8.5
)
```
#tribble is a quick way to enter data and organize it.  

### 10.3: Tibbles vs data.frame

Enter your code chunks for Section 10.2 here.
```{r}
tibble(
  a=lubridate::now() +runif(1e3)*86400,
  b=lubridate::today()+runif(1e3)*30,
  c=1:1e3,
  d=runif(1e3),
  e=sample(letters, 1e3, replace = TRUE)
)
```
#The tibble only shows 10 records or rows, and therefore is easier to view on the screen.  It also makes sure that all the columns appear on the screen.

```{r}
nycflights13::flights %>%
  print(n=10, width=Inf)
```
#by using the print function, we forced the data frame to print 10 rows at a time, but include all columns. 

```{r}
as.tibble(iris, tibble.print_max=n, tibble.print_min=m)
as.tibble(iris, tibble.print_min=Inf)
```

```{r}
nycflights13::flights %>%
  View()
```
#the last code block uses the viewer which creates a separate scrollable table.  
```{r}
df <- tibble(
  x=runif(5),
  y=rnorm(5)
)
#by name
df$x
df[["x"]]
#extract by posit
df[[1]]
```
#by using $ and [[ we can pull single variables out of a data frame. 
```{r}
df <- tibble(
  x=runif(5),
  y=rnorm(5)
)
df%>% .$x
df%>% .[["x"]]
```
#Using . will allow us to pull variables in a pipe.
Describe what each chunk code does. 



### 10.4: Not required

#### Section 10.5 Questions

Answer the questions *completely.* Use code chunks, text, or both, as necessary.

**1:** How can you tell if an object is a tibble? (Hint: try printing `mtcars`, which is a regular data frame). Identify at least two ways to tell if an object is a tibble. *Hint:* What does `as_tibble()` do? What does `class()` do? What does `str()` do?
```{r}
mtcars
```


```{r}
as.tibble(mtcars)
```
#Tibble is a more simple structure than the regular data frame.  Since tibble doesn't create row names, and mtcars has row names, I would say it is not a tibble.  Tibbles also default to printing only the first 10 rows and all the columns that fit on the screen.  Tibbles will not do partial matching whereas data.frames will.  
#as.tibble displays a data.frame as a tibble, with all the restrictions and simplicity.  class() helps define what attributes an object has, whereas str() displays the internal structure of the object.

**2:** Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?

```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]
```
```{r}
tb <- tibble(abc=1, xyz="a")
tb$"xyz"
tb[, "xyz"]
tb[, c("abc", "xyz")]
```
#The data.frame automatically converts xyz to factors x, y and z whereas tibble leaves xyz as a string of characters.  One problem with data.frame is that it ends up returning x="a" as well as xyz="a".  This may cause problems.  In the tibble, xyz="a" whereas x is still undefined.     

## Chapter 11: Importing data

Read [R4ds Chapter 11: Data Import](https://r4ds.had.co.nz/data-import.html), sections 1, 2, and 5.

### 11.1 Introduction

Nothing to do here unless you took a break and need to reload `tidyverse`.
```{r}
library(tidyverse)
```


### 11.2 Getting started.

Do *not* run the first code chunk of this section, which begins with `heights <- read_csv("data/heights.csv")`. You do not have that data file so the code will not run.

Enter and run the remaining chunks in this section.
```{r}
read_csv("a,b,c
1,2,3
4,5,6")
```
#Hmm, that didn't work.  Oh. Had to reload tidyverse.  Working fine now.  This example used the first line, "a,b,c" as the column names.  The following lines filled in as data. 

```{r}
read_csv("The first line of metadata
         The second line of metadata
         x,y,z
         1,2,3", skip=2)

read_csv("#A comment I want to skiperadoony
         x,y,z
         1,2,3", comment ="#")
```
#The first example skips two lines and begins creating the columns on the third line, which is great since the first two lines are metadata.  In the second example, the comment, which begins with "#" is removed with the comment="#" line. 
```{r}
read_csv("1,2,3\n4,5,6", col_names = FALSE)
```
#Since the data has no column names, col_names=FALSE instructs read_csv to create them using X1, X2, etc.
```{r}
read_csv("1,2,3\n4,5,6", col_names = c("x", "y", "z"))
```
#In this example, I used a character vector to name each column.
```{r}
read_csv("a,b,c\n1,2,.", na=".")
```
#this example shows how to insert a NA into the tibble using "."


#### 11.2 Questions

**1:** What function would you use to read a file where fields were separated with "|"?
#read_delim should be able to determine the field separation.


**2:** (This question is modified from the text.) Finish the two lines of `read_delim` code so that the first one would read a comma-separated file and the second would read a tab-separated file. You only need to worry about the delimiter. Do not worry about other arguments. Replace the dots in each line with the rest of your code. 

# Comma-separated
`file <- read_delim("file.csv", ...)`
```{r}
`file <- read_csv("file.csv",...)`
```


# Tab-separated
`file <- read_delim("file.csv", ...)`
```{r}
`file <- read_tsv("file...")
```


**3:** What are the two most important arguments to `read_fwf()`? Why?
#the most important arguments to read_fwf are col_positions and fwf_width because the data does not contain any deliminator information. The read_fwf function must be provided with the column positions and the width of each column in order to accurately organize the data into a table. 

**4:** Skip this question


**5: ** Identify what is wrong with each of the following inline CSV files. What happens when you run the code?

```{r}
read_csv("a,b\n1,2,3\n4,5,6")
```
#1. There are only two column headers listed, a,b, but there are three sets of variables per line.  Therefore, the number of columns doesn't match.  
 
```{r}
read_csv("a,b,c\n1,2\n1,2,3,4")

```
#2.The data has three columns, but the first row of data is missing a value for C and the second row of data has an extra column of data.  The graph is displayed, but the fourth value for row two is discarded.
```{r}
read_csv("a,b\n\"1")
```
#3. This code attempts to create a table with two columns, but then the first row has no data and a second forward slash. read_csv does its best and creates a table with one row with an a value of 1. 
```{r}
read_csv("a,b\n1,2\na,b")
```
#4. This table has column headings but then contains numeric and text data in the same columns. This seems like it might be problematic. Also, I'm wondering if the \na,b is supposed to show a NA value, like \n.,b"?", na="."?  
```{r}
read_csv("a;b\n1;3")
```
#5. Assuming that we want columns a, b and values of 1, 3, the code should contain commas, not semicolons. The use of semicolons creates one column that is character based.


### 11.3 and 11.4: Not required

### 11.5: Writing to a file


Just read this section. You may find it helpful in the future to save a data file to your hard drive. It is basically the same format as reading a file, except that you must specify the data object to save, in addition to the path and file name.

### 11.6 Not required

## Chapter 18: Pipes

Read [R4ds Chapter 18: Pipes](https://r4ds.had.co.nz/pipes.html), sections 1-3.
```{r}
msleep %>% 
  select(bodywt, brainwt, vore) %>%
  drop_na() %>%
  ggplot()+
  geom_point(aes(x=bodywt, y=brainwt, color=vore))
```

Nothing to do otherwise for this chapter. Is this easy or what?

**Note:** Trying using pipes for all of the remaining examples. That will help you understand them.

## Chapter 12: Tidy Data

Read [R4ds Chapter 12: Tidy Data](https://r4ds.had.co.nz/tidy-data.html), sections 1-3, 7. 

### 12.1 Introduction

Nothing to do here unless you took a break and need to reload the `tidyverse.`

### 12.2 Tidy data

Study Figure 12.1 and relate the diagram to the three rules listed just above them. Relate that back to the example I gave you in the notes. Bear this in mind as you make data tidy in the second part of this assignment.
#in the beetle example in the notes, Each variable needs its own column. In the case of plot and sample, that is correct, but species is broken into 5 different columns instead of one that represents how many appeared in the sample. Each observation must have its own row, which is correct in the example, and each value should have its own cell.

You do not have to run any of the examples in this section.

### 12.3

Read and run the examples through section 12.3.1 (gathering), including the example with `left_join()`. We'll cover joins later.
```{r}
table4a
```
```{r}
 table4a %>%
  gather(`1999`, `2000`, key="year", value="cases")
```
```{r}
table4b %>%
  gather(`1999`, `2000`, key="year", value="population")
```

```{r}
tidy4a <- table4a %>%
  gather(`1999`, `2000`, key = "year", value = "cases")
tidy4b <- table4b %>%
  gather(`1999`,  `2000`,  key="year",  value="population")
left_join(tidy4a, tidy4b)
```
```{r}
table2 %>%
  spread(key=type, value=count)
```

#### 12.3 Questions

**2:** Why does this code fail? Fix it so it works.

```{r}
table4a %>% 
  gather(1999, 2000, key = "year", value = "cases")
#> Error in inds_combine(.vars, ind_list): Position must be between 0 and n
```
#The years are non-syntatic names, so they need to be enclosed in backtick marks.
```{r}
table4a %>%
  gather(`1999`, `2000`, key = "year", value = "cases")
```


That is all for Chapter 12. On to the last chapter.


## Chapter 5: Data transformation

Read [R4ds Chapter 5: Data Transformation](https://r4ds.had.co.nz/transform.html), sections 1-4.

Time to [get small.](https://www.youtube.com/watch?v=GOrdzCHnpw4) 

### 5.1: Introduction

Load the necessary libraries. As usual, type the examples into and run the code chunks.
```{r}
library(nycflights13)
library(tidyverse)
```
```{r}
flights
```

### 5.2: Filter rows with `filter()`
```{r}
filter(flights, month==1, day==1)
```
```{r}
jan1 <- filter(flights, month==1, day==1)
jan1
```
```{r}
(dec25 <- filter(flights, month==12, day==25))
```
```{r}
filter(flights, month=1)
```

```{r}
sqrt(2)^2==2
1/49*49==1
```
```{r}
near(sqrt(2)^2, 2)
near(1/49*49, 1)
```
```{r}
filter(flights, month==11|month==12)
```
```{r}
nov_dec <- filter(flights, month %in% c(11,12))
nov_dec
```
```{r}
filter(flights, !(arr_delay>120 | dep_delay>120))
filter(flights, arr_delay<=120, dep_delay<=120)
```
```{r}
NA>5
10==NA
NA+10
NA/2
```
```{r}
NA==NA
```
```{r}
x <- NA
y <- NA
x==y
```
```{r}
is.na(x)
```
```{r}
df <- tibble(x=c(1, NA, 3))
filter(df, x>1)
filter(df, is.na(x) |x>1)
```





Study Figure 5.1 carefully. Once you learn the `&`, `|`, and `!` logic, you will find them to be very powerful tools.


#### 5.2 Questions

**1.1:** Find all flights with a delay of 2 hours or more.
```{r}

```



**1.2:** Flew to Houston (IAH or HOU)


**1.3:** Were operated by United (UA), American (AA), or Delta (DL).


**1.4:** Departed in summer (July, August, and September).


**1.5:** Arrived more than two hours late, but didn’t leave late.

**1.6:** Were delayed by at least an hour, but made up over 30 minutes in flight. This is a tricky one. Do your best.


**1.7:** Departed between midnight and 6am (inclusive)


**2:** Another useful dplyr filtering helper is `between()`. What does it do? Can you use it to simplify the code needed to answer the previous challenges?


**3:** How many flights have a missing dep_time? What other variables are missing? What might these rows represent?


**4:** Why is `NA ^ 0` not missing? Why is `NA | TRUE` not missing? Why is `FALSE & NA` not missing? Can you figure out the general rule? (`NA * 0` is a tricky counterexample!)

**Note:** For some context, see [this thread](https://blog.revolutionanalytics.com/2016/07/understanding-na-in-r.html)


### 5.3 Arrange with `arrange()`


#### 5.3 Questions

**1:** How could you use `arrange()` to sort all missing values to the start? (Hint: use is.na()). **Note:** This one should still have the earliest departure dates after the `NA`s. *Hint:* What does `desc()` do?


**2:** Sort flights to find the most delayed flights. Find the flights that left earliest. 

This question is asking for the flights that were most delayed (left latest after scheduled departure time) and least delayed (left ahead of scheduled time).


**3:** Sort flights to find the fastest flights. Interpret fastest to mean shortest time in the air.


*Optional challenge:* fastest flight could refer to fastest air speed. Speed is measured in miles per hour but time is minutes. Arrange the data by fastest air speed.


**4:** Which flights travelled the longest? Which travelled the shortest?


### 5.4 Select columns with `select()`

#### 5.4 Questions

**1:** Brainstorm as many ways as possible to select `dep_time`, `dep_delay`, `arr_time`, and `arr_delay` from flights. Find at least three ways.


**2:** What happens if you include the name of a variable multiple times in a `select()` call?



**3:** What does the `one_of()` function do? Why might it be helpful in conjunction with this vector?

`vars <- c("year", "month", "day", "dep_delay", "arr_delay")`



**4:** Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

`select(flights, contains("TIME"))`



