---
title: "Chronister_08.Rmd"
author: Andrew Chronister"
output: html_notebook
date: "`r format(Sys.time(), '%d %B %Y')`"
editor_options: 
  chunk_output_type: inline
---

***
###Load the Libraries
```{r}
library(tidyverse)
library(lubridate)
```

***
### Part 1: Play date
#setting the variables/values
```{r}
birth_date <- "08/23/1968"

# Elizabeth Garret Anderson was the first female physician
anderson <- "9 June 1836"

# Mary Anning discovered the first complete Plesiosaur fossil, 
# and other important Jurassic fossils
anning <-  "1799/05/21"

# Alice Augusta Ball was a chemist who developed the
# first injectable treatment for leprosy.
ball <- "July 24, 1892"

# Sylvia Earle was the first chieft scientist at NOAA
earle <- "August, 30 (1935)"

# Rosalind Franklin used x-ray diffraction to 
# discover DNA structure. Watson and Crick claimed credit.
franklin <- "25th of July, 1920"

# Ester Lederberg was the first to discover that viruses 
# infect bacteria. Led to discovery of lambda bacteriophage
lederberg <- "December 18, 1922"

# Barbara McClintock discovered transposable elements (jumping genes),
# and developed staining techniques to study chromosomes
mcclintock <- "16061902"

# Martin Luther King's birthday is coming up. Let's remember a dream
mlk_birth <- "1/15/29"
mlk_dream <- "1963 August 08"
```
#1. Convert to POSIXct format using lubridate functions.
```{r}
chronister <- mdy(birth_date)
anderson <- dmy(anderson)
anning <- ymd(anning)
ball <- mdy(ball)
earle <- mdy(earle)
franklin <- dmy(franklin)
lederberg <- mdy(lederberg)
mcclintock <- dmy(mcclintock)
mlk_birth <- mdy(mlk_birth)
mlk_dream <- ymd(mlk_dream)
chronister
anderson
anning
ball
earle
franklin
lederberg
mcclintock
mlk_birth
mlk_dream

```
#2. Check the dates to make sure they are correct.  
#All the dates appear to be good, except MLK's birthday, which shows up as 2029 because of the 1968 rule.I believe I can fix it.
```{r}
mlk_birth <- update(mlk_birth, year = 1929)
mlk_birth
```
#That looks more like it.  The mlk_birth variable has been updated to reflect the correct birth year.  Good thing I'm not 8 months older or I would have to modify my birth year as well.

#3. What day of the week was I born on?
```{r}
wday(chronister, label = TRUE, abbr = FALSE)
```
#It looks like I was born on a Friday.  

#4. What day of the week will birthday be this year? (The way I understand this question is that the code should update itself depending on what year it is, so that in 2020, it will grab the current year and use that data.  That is how I'm going to approach this. If one wanted to be able to look up what weekday the birthday would be five years in the future, there would need to be an edit made in order to enter what year is to be used.)
```{r}
what_yr_is_it <- year(today())
update(chronister, year = what_yr_is_it)
wday(chronister, label = TRUE, abbr = FALSE)
```
#It isn't very pretty, but it appears to be right.  Probably should have done the two steps in pipes.  

#5. What day of the week will my birthday be celebrated on in 2525? :D
```{r}
chronister %>%
  update(year = 2525) %>%
  wday(label = TRUE, abbr = FALSE)
```
#Looks like "Chronister Day" will be on a Thursday in 2525.  Don't know what they would be celebrating though.  

#6. How many days between the birth of Anderson and Ball.
```{r}
anderson - ball
```
#It can't be that easy... can it? Wow.

#7. Duration between MLK birth and MLK dream.
```{r}
mlk_duration <- mlk_dream - mlk_birth
mlk_duration <- as.duration(mlk_duration)
mlk_duration

```
#Looks like MLK was 34 years old when he gave his famous speech.

#Enter eight birthdates into vector.
```{r}
birth_vector <- c("1968-08-23","1836-06-09","1799-05-21","1892-07-24","1935-08-30","1920-07-25","1922-12-18","1902-06-16","1929-01-15","1963-08-08")
names(birth_vector) <- c("chronister", "anderson", "anning", "ball", "earle", "franklin", "lederberg", "mcclintock", "mlk_birth", "mlk_dream")
```
#Which birthdates are leap years?

```{r}
YearsOnly <- year(birth_vector)
LeapYears <- leap_year(YearsOnly)
Leaps <- YearsOnly[LeapYears]
Leaps

```
#Looks like my birth year was a leap year, so there were four in the list.  I think I'm supposed to leave the names in there, but could have removed them with the unname(birth_vector).

```{r}

```





