---
title: "Chronister_08.Rmd"
author: Andrew Chronister"
output: html_notebook
date: "`r format(Sys.time(), '%d %B %Y')`"
editor_options: 
  chunk_output_type: inline
---

***
###Load the Libraries
```{r}
library(tidyverse)
library(lubridate)
```

***
### Part 1: Play date
#setting the variables/values
```{r}
birth_date <- "08/23/1968"

# Elizabeth Garret Anderson was the first female physician
anderson <- "9 June 1836"

# Mary Anning discovered the first complete Plesiosaur fossil, 
# and other important Jurassic fossils
anning <-  "1799/05/21"

# Alice Augusta Ball was a chemist who developed the
# first injectable treatment for leprosy.
ball <- "July 24, 1892"

# Sylvia Earle was the first chieft scientist at NOAA
earle <- "August, 30 (1935)"

# Rosalind Franklin used x-ray diffraction to 
# discover DNA structure. Watson and Crick claimed credit.
franklin <- "25th of July, 1920"

# Ester Lederberg was the first to discover that viruses 
# infect bacteria. Led to discovery of lambda bacteriophage
lederberg <- "December 18, 1922"

# Barbara McClintock discovered transposable elements (jumping genes),
# and developed staining techniques to study chromosomes
mcclintock <- "16061902"

# Martin Luther King's birthday is coming up. Let's remember a dream
mlk_birth <- "1/15/29"
mlk_dream <- "1963 August 08"
```
#1. Convert to POSIXct format using lubridate functions.
```{r}
chronister <- mdy(birth_date)
anderson <- dmy(anderson)
anning <- ymd(anning)
ball <- mdy(ball)
earle <- mdy(earle)
franklin <- dmy(franklin)
lederberg <- mdy(lederberg)
mcclintock <- dmy(mcclintock)
mlk_birth <- mdy(mlk_birth)
mlk_dream <- ymd(mlk_dream)
chronister
anderson
anning
ball
earle
franklin
lederberg
mcclintock
mlk_birth
mlk_dream

```
#2. Check the dates to make sure they are correct.  
#All the dates appear to be good, except MLK's birthday, which shows up as 2029 because of the 1968 rule.I believe I can fix it.
```{r}
mlk_birth <- update(mlk_birth, year = 1929)
mlk_birth
```
#That looks more like it.  The mlk_birth variable has been updated to reflect the correct birth year.  Good thing I'm not 8 months older or I would have to modify my birth year as well.

#3. What day of the week was I born on?
```{r}
wday(chronister, label = TRUE, abbr = FALSE)
```
#It looks like I was born on a Friday.  

#4. What day of the week will birthday be this year?

```{r}

```

