---
title: "HW 07 Part 1: Data Visualization"
author: "Andrew Chronister"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
```{r}

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(GGally)
library(patchwork)
```

#Setting the data filepath
```{r}
file_path <-"~/Documents/BI685Files/GitHubDocs/chronister_andrew/data"
sparrows_raw <- read_tsv(file.path(file_path, "sparrows.txt"))
View(sparrows_raw)
sparrows_raw
```
```{r}
sparrows <- sparrows_raw %>%
  mutate(box_group = "Sparrows",
         row_order = 1:nrow(.))
View(sparrows)
```
#Figure 2: Outliers 
```{r}
plot_box <- sparrows %>%
  ggplot(aes(x = box_group,
             y = wingcrd))+
  geom_boxplot(width = 0.25)+
  labs(x = NULL,
       y = "Wing length (mm)")
```
#Create Cleveland plot
```{r}
plot_cleveland <- sparrows %>%
  ggplot(aes(x = wingcrd,
             y = row_order))+
  geom_point()+
  labs(x = "Wing length (mm)",
       y = "Order of the data")
```
#Store plots into variables...
#Plot both plots side by side.
```{r}
plot_box + plot_cleveland
```
#Add Coordinate Flip so scales line up
```{r}
plot_box + plot_cleveland + coord_flip()
```
#Create Culmen variable
```{r}
p_culmen <-  sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = culmen)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
p_culmen
```
#Create nalospi variable
```{r}
p_nalospi <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = nalospi)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
p_nalospi
```
#Create wt variable
```{r}
p_wt <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wt)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
p_wt
  
```
#Create wingcrd variable (I'm using copy/paste on the three lines of theme because I'm lazy...)
```{r}
p_wingcrd <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wingcrd)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
p_wingcrd
```
#Create tarsus variable
```{r}
p_tarsus <- sparrows %>%
  ggplot(aes(y = row_order))+
  geom_point(aes(x = tarsus))+
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
p_tarsus
```
#Create head variable
```{r}
p_head <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = head)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
p_head
```
#Now, put them all together in one figure. 
```{r}
p_culmen <-  sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = culmen)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
p_nalospi <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = nalospi)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
p_wt <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wt)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
p_wingcrd <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wingcrd)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
p_tarsus <- sparrows %>%
  ggplot(aes(y = row_order))+
  geom_point(aes(x = tarsus))+
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
p_head <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = head)) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
p_culmen + p_nalospi + p_wt + p_wingcrd + p_tarsus + p_head + plot_layout(ncol = 3, nrow = 2)
```
#It looks like there is one possible outlier in nalospi, one in wingcrd, one in tarsus and two in head. All of these would need further consideration, just like the wingcrd example, to determine if the value is practical or an error. 

####Figure 5
#First code block to re-create the histogram
```{r}
h1 <- sparrows %>%
  filter(Month %in% 6:8) %>%
  ggplot() +
  geom_histogram(aes(x = wt),
                 binwidth = 0.5,
                 boundary = 0,
                 closed = "right",
                 color = "black") +
  scale_x_continuous(breaks = seq(14,28,by = 2))
h1
```
#Filter all but June, July and August, etc.
```{r}
h2 <- sparrows %>%
  filter(Month %in% 6:8) %>%
  mutate(Month = case_when(
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August")) %>%
  ggplot() +
  geom_histogram(aes(x = wt),
                 binwidth = 0.5,
                 boundary = 0,
                 color = "black") +
  scale_x_continuous(breaks = seq(14,28,by = 2)) +
  facet_wrap(~Month, ncol = 1)
h2
```
#Now assemble into one figure.
```{r}
h1 + h2
```
#Grad/Honors density plot.
#Left panel playground(where I'll figure out how to do it)
```{r}
dp_1 <- sparrows %>%
  filter(Month %in% 6:8) %>%
  ggplot() +
  geom_density(aes(x = wt)) +
  scale_x_continuous(breaks = seq(14,28, by = 2))
dp_1
```
#Ok. That looks similar enough to the example in the notes.  Now, on to the filled density plot trinity.  Breakthrough. That is kinda pretty...
```{r}
dp_2 <- sparrows %>%
  filter(Month %in% 6:8)%>%
  mutate(Month = case_when(
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August")) %>%
  ggplot()+
  geom_density(aes(x = wt,
                   fill = Month),
               alpha = 0.42)+
  scale_x_continuous(breaks = seq(14, 28, by = 2))
dp_2
```
#Now simply put the two together. 
```{r}
dp_1 + dp_2 + plot_layout(nrow = 1)
```
#Create ggscatmat plot of data using first 7 columns
```{r}
sparrows %>% ggscatmat(columns = 1:7)
```
#Tis a thing of beauty, and really, really easy.  

#Figure 11: Interactions
```{r}
month_order <-  c("May", "June", "July", "August", "September")
sparrows %>%
  filter(Month %in% 5:9,
         Sex != 0) %>%
  mutate(Month = case_when(
    Month == 5 ~ "May",
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August",
    Month == 9 ~ "September"),
    Month = factor(Month, levels = month_order, ordered = TRUE),
    Sex = ifelse(Sex == 4, "Male", "Female")) %>%
  ggplot(aes(x = wingcrd,
             y = wt))+
  geom_point()+
  geom_smooth(method = lm, se = FALSE)+
  facet_grid(Sex ~ Month)
```
#### Part 2:Aegla Crab-fest 
#import the raw data
```{r}
aegla_raw <- read_csv(file.path(file_path, "aegla_crabs.csv"))
aegla_raw
```
#Select to remove certain columns
```{r}
aegla_select <- select(aegla_raw, -Site, -AT, -WT, -Shrimp, -Aeglam, -AeglaFe, -AeglaFo)
aegla_select
```
#Use drop_na to drop row with NAs
```{r}
aegla_dropna <- aegla_select %>%
  drop_na(4:12)
aegla_dropna
```
#Create the row_order dummy variable
```{r}
aegla_clean <- aegla_dropna %>%
  mutate(box_group = "Crabs",
         row_order = 1:nrow(.))
aegla_clean
```
#Trying the more efficient pipe method.
```{r}
aegla <- aegla_raw %>%
  select(-c("Site", "AT", "WT", "Shrimp", "Aeglam", "AeglaFe", "AeglaFo"))%>%
  drop_na(c("N"))%>%
  mutate(box_group = "Crabs",
         row_order = 1:nrow(.))
aegla
```
#Explore the data
```{r}
p_width <- aegla %>%
  ggplot(aes(x=row_order))+
  geom_point(aes(y=Width))
p_depth <- aegla %>%
  ggplot(aes(x=row_order))+
  geom_point(aes(y=Depth))
p_flow <- aegla %>%
  ggplot(aes(x=row_order))+
  geom_point(aes(y=Flow))
p_ph <- aegla %>%
  ggplot(aes(x=row_order))+
  geom_point(aes(y=pH))
p_tds <- aegla %>%
  ggplot(aes(x=row_order))+
  geom_point(aes(y=TDS))
p_cond<- aegla %>%
  ggplot(aes(x=row_order))+
  geom_point(aes(y=Cond))
p_n <- aegla %>%
  ggplot(aes(x=row_order))+
  geom_point(aes(y=N))
p_ni <- aegla %>%
  ggplot(aes(x=row_order))+
  geom_point(aes(y=Ni))
p_nt <- aegla %>%
  ggplot(aes(x=row_order))+
  geom_point(aes(y=Nt))
p_phos <- aegla %>%
  ggplot(aes(x=row_order))+
  geom_point(aes(y=Phos))
p_mg <- aegla %>%
  ggplot(aes(x=row_order))+
  geom_point(aes(y=Mg))
p_ca <- aegla %>%
  ggplot(aes(x=row_order))+
  geom_point(aes(y=Ca))
p_crab <- aegla %>%
  ggplot(aes(x=row_order))+
  geom_point(aes(y=Crab))
p_width + p_depth + p_flow 

```

```{r}
p_ph + p_tds + p_cond  
```
```{r}
p_n + p_ni + p_nt 
```

```{r}
p_phos + p_mg + p_ca + p_crab
```
```{r}
ggplot(aegla, aes(x=row_order, y=Phos))+
  geom_boxplot()+
  geom_point()
```
#The outliers I found are sample 1, where the Phosphate levels are almost 4 times higher than the next highest value, sample 14 has low pH at 6.25, with the next lowest measurement being 6.75 and 0 waterflow, and sample 22 was taken at a width of 12.5 (meters?) and a depth of 5 meters.  The width measurement is not much of an outlier, but the depth measurement is twice as deep as the next highest value.  I would contend that the real outlier is the phosphate level in sample 1.  

#Plot four variables with a Cleveland plot. It just so happened that I made plots for all of the variables... 
```{r}
p_width + p_depth + p_cond + p_crab + plot_layout(nrow = 2, ncol = 2)
```
#Make three histograms
```{r}
ha1 <- ggplot(data = aegla)+
  geom_histogram(aes(x = N),
                 binwidth = .0025,
                 boundary = 0,
                 closed = "right",
                 color = "black")
ha1



```
```{r}
ha2 <- ggplot(data = aegla)+
  geom_histogram(aes(x=Phos),
                 binwidth = .005,
                 boundary = 0,
                 closed = "right",
                 color = "blue")
ha2
```
```{r}
ha3 <-  ggplot(aegla)+
  geom_histogram(aes(x=pH),
                 binwidth = .005,
                 boundary = 0,
                 closed = "right",
                 color = "red")
ha3
```
#Create density plots
```{r}
hdp1 <- ggplot(aegla)+
  geom_density(aes(x=N))
hdp1

```
```{r}
hdp2 <- ggplot(aegla)+
  geom_density(aes(x=Phos))
hdp2
```
```{r}
hdp3 <- ggplot(aegla)+
  geom_density(aes(x=pH))
hdp3
```
#Use patchwork to produce 2 col matrix
```{r}
ha1 + hdp1 + ha2 + hdp2 + ha3 + hdp3 + plot_layout(ncol=2, nrow = 3)
```


