---
title: "HW 09: Strings"
author: "Andrew Chronister"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
***
###Install and load the libraries
```{r}
install.packages("ggthemes")
install.packages("RColorBrewer")
install.packages("viridis")
library(tidyverse)
library(htmlwidgets)

```
***
###Define the file.path
```{r}
file_path <- "~/Documents/BI685Files/GitHubDocs/chronister_andrew/data"
```


***
###Part 1:DNA Strings
***
####Read in the data
```{r}
(gliadin_raw <- read_lines(file.path(file_path, "beta_gliadin.txt"), skip = 1))
```
***
###"Tidy" and "Wrangle" the data
#Getting rid of the spaces, digits and tabs.
```{r}
gliadin_interm <- gliadin_raw%>%
  str_replace_all(" ", "")%>%
  str_replace_all("\\d", "")%>%
  str_replace_all("\\t", "")
gliadin_interm
```
#Collapse the strings into just one string.
```{r}
gliadin_seq <- gliadin_interm%>%
  str_c(collapse = "")
gliadin_seq
```
#Convert case to upper.
```{r}
gliadin_seq <- gliadin_seq%>%
  str_to_upper()
gliadin_seq
str_length(gliadin_seq)
```
#The DNA sequence is 3310 characters long. I guess I got it right.  

***
###Sequence analysis: find the promoter region
#Use string count to find non ACGT letters.
```{r}
str_count(gliadin_seq, "[^ACGT]")
```
#I came up with 0 letters that are not ACGT.

#Count possible Start codons. Count possible Stop codons.
```{r}
str_count(gliadin_seq, "ATG")
str_count(gliadin_seq, "TGA|TTA")
```
#Assign CAAT and TATA box variables
```{r}
caat_box <- c("CCAAT")
tata_box <- c("CTATAA")
str_count(gliadin_seq, caat_box)
str_count(gliadin_seq, tata_box)
```
#Create CAAT sequence searrch string.
```{r}
caat_seq <- "CCAAT.{100,150}ATG"
str_count(gliadin_seq, caat_seq)
```
#The CAAT sequence occurs twice in those conditions. I got hung up on this for a few minutes because a space worked its way in between the "(" and the "ATC".  I kept coming up with 0 in the results.  
#Now for the TATA box extraction
```{r}
caat_seqs_all <- str_extract_all(gliadin_seq, caat_seq, simplify = TRUE)
caat_seqs_all
str_count(caat_seqs_all)

```
#Use str_which to find which seq has tatabox
```{r}
tata_seq <- "CTATAA.{50,100}ATG"
caat_tata <- str_which(caat_seqs_all, tata_seq)
promotor <- str_subset(caat_seqs_all, tata_seq)
promotor
str_length(promotor)
```

***
###Part 2: Ants!
***
###Load the Libraries and define the file path. (packages are already installed, but if they weren't, I'd be doing this.)
```{r}
install.packages("ggthemes")
install.packages("RColorBrewer")
install.packages("viridis")
library(tidyverse)
library(htmlwidgets)
file_path <- "~/Documents/BI685Files/GitHubDocs/chronister_andrew/data"
```
***
###Read in the data.
```{r}
ants_raw <- read_csv(file.path(file_path, "ants.csv"))
ants_raw
```

***
###Tidy the data and wrangle it
```{r}
ants_data <- ants_raw %>%
  select(-c(season, seas.code, trt, plot, sample))%>%
  gather(key = species, value = abundance, matches("\\w\\.\\w", ignore.case = FALSE))
ants_data
```
#Removing the seas.code row makes the gather much easier.  

#The hardest part maybe...
```{r}
ant_abundance <- ants_data%>%
  mutate(
    species = str_replace(
      species,
      pattern = "(^[A-Z]*\\w{3})\\w+\\.(\\w{4})\\w+",
    replacement = "\\1\\2"),
    year = factor(year, levels = c("2010", "2011"), labels = c("2010", "2011"), ordered = TRUE))%>%
  group_by(year, species)%>%
  summarize(total = sum(abundance, na.rm=TRUE))
ant_abundance
```
#Sheesh.  My brain hurts.  

***
###Plot the data
```{r}
ant_abundance%>%
  ggplot(aes(x = reorder(species, total), y = total, group = species))+
  geom_line(color = "gray")+
  geom_point(aes(shape = year), size = 2)+
  coord_flip()+
  labs(x = NULL, y = "Total Abundance", shape = "Year")+
  theme_minimal()
```
#So in the slope graph, you are simply adding the points and text to the ends of a line graph. Interesting.

***
###Part 3:Featuring Phlox Flowers
***
###Install the libraries and set file_path. 
```{r}
library(tidyverse)
library(htmlwidgets)
library(Polychrome)
file_path <- "~/Documents/BI685Files/GitHubDocs/chronister_andrew/data"
```
***
###Read in the data
```{r}
(flowers_raw <- read.csv(file.path(file_path, "flower_size.csv")))

```
***
###Tidy up the data
```{r}
flowers_tidy <- flowers_raw%>%
  filter(!is.na(Flower.number))%>%
  select(species = Species, cor_length = Corolla.length..cm., throat_width = Corolla.width.throat..cm.)
flowers_tidy
```
#Separate the genus and species.
```{r}
flowers_tidy <- flowers_tidy%>%
  separate(species, into = c("genus", "species"), sep = " ", extra = "drop")
flowers_tidy
```
#Group data and summarize the means
```{r}
flowers_tidy <- flowers_tidy%>%
  group_by(genus, species)%>%
  summarize(mean_length = mean(cor_length), mean_width = mean(throat_width), max_length = max(cor_length), max_width = max(throat_width))
flowers_tidy
```

***
###Import and wrangle second data set
```{r}
(pollinators_raw <- read_csv(file.path(file_path, "flower_pollinators.csv")))
```
```{r}
pollinators <- pollinators_raw %>%
  select(-c(Color, starts_with("Source"))) %>%
  filter(!is.na(Pollinator)) %>%
  separate(Species, into = c("genus", "species"), sep = " ", extra = "drop")
pollinators
```

#separate out the major polinator
```{r}
pollinators <- pollinators%>%
  separate(Pollinator, into = "Syndrome", sep = " |,", extra = "drop")
pollinators
```

#Join the datasets
```{r}
phlox <- inner_join(flowers_tidy, pollinators, by = c("genus", "species"))
phlox
```
#Plotting
```{r}
phlox%>%
  ggplot(aes(x=mean_length, y=mean_width, color=Syndrome))+
  geom_point()
```
#combine the bees and the beetles
```{r}
phlox <- phlox%>%
  mutate(
    Syndrome = str_replace(Syndrome, "bee$", "bees"),
    Syndrome = str_replace(Syndrome, "beetle$", "beetles"))

```
#Replot
```{r}
phlox%>%
  ggplot(aes(x=mean_length, y=mean_width, color=Syndrome))+
  geom_point()
```
#Run the code
```{r}
phlox%>%
  group_by(Syndrome)%>%
  count(Syndrome)
```
#Lump beetles and flies together
```{r}
phlox <- phlox%>%
  mutate(Syndrome = str_replace(Syndrome, "bees|beetles", "other insects"))
phlox
```
#Plot again
```{r}
phlox%>%
  ggplot(aes(x=mean_length, y=mean_width, color=Syndrome))+
  geom_point()
```
#create a new column called syndrome group
```{r}
phlox <- phlox%>%
  mutate(
    syndrome_group = case_when(Syndrome == "autogamous" ~ "Autogamous",
                               Syndrome == "bat" ~ "Mammal",
                               Syndrome == "hummingbird" ~ "Bird", 
                               TRUE ~ "Insect" )
  )
phlox

```
#Replot
```{r}
phlox%>%
  ggplot(aes(x=mean_length, y=mean_width, color=syndrome_group))+
  geom_point()
```
***
###Improving the plots
#using brewer palaettes
```{r}
phlox%>%
  ggplot()+
  geom_point(aes(x=mean_length,
                 y=mean_width,
                 color=syndrome_group))+
  scale_color_brewer(type = "qual", palette = "Dark2")
```
#experimenting with colors
```{r}
phlox%>%
  ggplot()+
  geom_point(aes(x=mean_length,
                 y=mean_width,
                 color=syndrome_group))+
  scale_color_brewer(type = "div", palette = "BrBG")
```
```{r}
phlox%>%
  ggplot()+
  geom_point(aes(x=mean_length,
                 y=mean_width,
                 color=syndrome_group))+
  scale_color_brewer(type = "seq", palette = "Greens")
```
```{r}
phlox%>%
  ggplot()+
  geom_point(aes(x=mean_length,
                 y=mean_width,
                 color=syndrome_group))+
  scale_color_viridis_d(option = "magma")
```
```{r}
phlox%>%
  ggplot()+
  geom_point(aes(x=mean_length,
                 y=mean_width,
                 color=syndrome_group))+
  scale_color_viridis_d(option = "inferno")
```
#The mammal color is really hard to see.  
```{r}
semo_palette <- c("#C8102E", "#DBE245", "#71C5E8", "#B07C57")
phlox%>%
  ggplot()+
  geom_point(aes(x=mean_length,
                 y=mean_width,
                 shape=syndrome_group,
                 fill=syndrome_group))+
  scale_shape_manual(values = c(22:25))+
  scale_fill_manual(values = semo_palette)
```
#Tufte plot
```{r}
library(ggthemes)
semo_palette <- c("#C8102E", "#DBE245", "#71C5E8", "#B07C57")
phlox%>%
  ggplot()+
  geom_point(aes(x=mean_length,
                 y=mean_width,
                 shape=syndrome_group,
                 fill=syndrome_group),
             size = 2)+
  labs(shape="Syndrome",
       fill="Syndrome",
       x="Mean Length (cm)",
       y="Mean Width (cm)")+
  theme_tufte()+
  scale_shape_manual(values=c(22:25))+
  scale_fill_brewer(palette = "Dark2")+
  theme(axis.text = element_text(size = 10),
        axis.title = element_text(size = 12))
```

***
###Part 4:Customize your plots
#Make four scatterplots with maxwidth as a function of maxlength
##THIS PLOT IS NOT ONE OF THE FOUR.  I'm just printing the most basic plot.
```{r}
phlox%>%
  ggplot(aes(x=max_length, y=max_width, color=syndrome_group))+
  geom_point()
```
#Plot1
```{r}
copper_dome <- c("#B07C57", "#7A6855", "#8A8F64", "#A4A9AD")
phlox%>%
  ggplot()+
  geom_point(aes(x=max_length,
                 y=max_width,
                 shape=syndrome_group,
                 fill=syndrome_group),
             size=2)+
  labs(shape="Syndrome",
       fill="Syndrome",
       x="Max Length (cm)",
       y="Max Length (cm)")+
  scale_shape_manual(values = c(22:25))+
  scale_fill_manual(values = copper_dome)
```
#Plot2
```{r}
phlox%>%
  ggplot()+
  geom_point(aes(x=max_length,
                 y=max_width,
                 shape=syndrome_group,
                 fill=syndrome_group))+
  labs(shape="Syndrome",
       fill="Syndrome",
       x="Max Length (cm)",
       y="Max Length (cm)")+
  theme_solarized()+
  scale_fill_brewer(palette = "Accent")+
  scale_shape_manual(values = c(21:24))
 
```
#Plot3
```{r}
phlox%>%
  ggplot()+
  geom_point(aes(x=max_length,
                 y=max_width,
                 shape=syndrome_group,
                 fill=syndrome_group))+
  labs(shape="Syndrome",
       fill="Syndrome",
       x="Max Length (cm)",
       y="Max Length (cm)")+
  theme_economist()+
  scale_shape_manual(values=c(21:24))
  
```
#Plot4
```{r}
phlox%>%
  ggplot()+
  geom_point(aes(x=max_length,
                 y=max_width,
                 shape=syndrome_group,
                 fill=syndrome_group),
             size = 3)+
  labs(shape="Syndrome",
       fill="Syndrome",
       x="Max Length (cm)",
       y="Max Length (cm)")+
  theme_fivethirtyeight()+
  scale_shape_manual(values = (21:24))+
  scale_fill_fivethirtyeight()
  
   
 
  
  
  
```
#Plot5, make it ugly
```{r}
phlox%>%
  ggplot()+
  geom_point(aes(x=max_length,
                 y=max_width,
                 shape=syndrome_group,
                 fill=syndrome_group),
             size = 3)+
  labs(shape="Syndrome",
       fill="Syndrome",
       x="Max Length (cm)",
       y="Max Length (cm)")+
  theme_wsj()+
 scale_shape_manual(values = (21:24))+
  scale_color_excel()
  
```
#I used excel colors.  Don't know if it gets uglier than that.  





