---
title: "HW10: Maps"
author: "Andrew Chronister"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

***
###Install packages
***
###Load libraries
```{r}
library(raster)
library(tidyverse)
library(ggthemes)
library(patchwork)
library(maptools)
library(ggmap)
library(marmap)
```
***
###Part 1: Outline maps
```{r}
usa <- map_data(map="state")
us_missouri <- map_data(map="state",
                        region="missouri")
us_other <- subset(x=usa,
                   subset=region %in% c("alabama",
                                        "florida",
                                        "mississippi",
                                        "north carolina"))
```

```{r}
str(usa)
```
```{r}
ggplot()+
  geom_polygon(data=us_other,
               aes(x=long,
                   y=lat,
                   group=group),
               fill = "gray")
```

```{r}
ggplot()+
  geom_polygon(data = us_other,
               aes(x=long,
                   y=lat,
                   group=group),
               fill="gray")+
  geom_polygon(data=us_missouri,
               aes(x=long,
                   y=lat,
                   group=group),
               fill="black")
```

```{r}
ggplot()+
  geom_polygon(data = us_other,
               aes(x=long,
                   y=lat,
                   group=group),
               fill="gray")+
  geom_polygon(data=us_missouri,
               aes(x=long,
                   y=lat,
                   group=group),
               fill="black")+
  geom_polygon(data = usa,
               aes(x=long,
                   y=lat,
                   group=group),
               fill=NA,
               color="black")
```

```{r}
us_map <- ggplot()+
  geom_polygon(data = us_other,
               aes(x=long,
                   y=lat,
                   group=group),
               fill="gray")+
  geom_polygon(data=us_missouri,
               aes(x=long,
                   y=lat,
                   group=group),
               fill="black")+
  geom_polygon(data=usa,
               aes(x=long,
                   y=lat,
                   group=group),
               fill=NA,
               color="black")+
  theme_map()+
  coord_fixed(1.3)
us_map
```
#Missouri map 
```{r}
missouri <- map_data("county", "missouri")
mo_scott <- subset(missouri, subregion %in% c("scott"))
mo_map <- ggplot()+
  geom_polygon(data=missouri, aes(x=long, y=lat, group=group), fill=NA, color="black")+
  geom_polygon(data=mo_scott, aes(x=long, y=lat, group=group), fill="black")+
  theme_map()+
  coord_fixed(1.3)
mo_map
```

```{r}
us_map + mo_map + plot_layout(ncol = 2, widths = c(1.5, 1))
```
#Outline Maps
#This should be interesting... I was born in Hawaii. It isn't part of the map_data, so I'll choose Utah, where I lived for two years and learned to ski.  I lived in Ogden in Weber county.  My initials are AKC. The counties I chose to put in gray are Kane and Carbon.   
```{r}
us_utah <- map_data(map="state",
                      region = "utah")
utah <- map_data("county", "utah")
ut_birth <- subset(utah, subregion %in% c("weber"))
ut_names <- subset(utah, subregion %in% c("carbon", "kane"))

akc_usa <- ggplot()+
  geom_polygon(data = us_utah,
               aes(x = long,
               y = lat,
               group = group),
               fill = "black")+
  geom_polygon(data = usa,
               aes(x = long,
               y = lat,
               group = group),
               fill = NA,
               color = "black")+
  theme_map()+
  coord_fixed(1.3)

akc_counties <- ggplot()+
  geom_polygon(data = ut_birth,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = "black")+
  geom_polygon(data = ut_names,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = "gray")+
  geom_polygon(data = utah,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = NA,
               color = "black")+
  theme_map()+
  coord_fixed(1.3)

akc_usa + akc_counties + plot_layout(ncol = 2, widths = c(1.5,1))
```
#The Hawaii Gambit... Born on Oahu, and have visited Hawaii (the big island) and Maui. I'll see if I can get that to work.  I love working with maps.    
```{r}
hawaii <- map_data("world", "USA:Hawaii")
hawaii_oahu <- map_data("world", "USA:Hawaii:Oahu")
other_islands <- map_data("world", c("USA:Hawaii:Hawaii", "USA:Hawaii:Maui"))

akc_usa <- ggplot()+
  geom_polygon(data = us_utah,
               aes(x = long,
               y = lat,
               group = group),
               fill = "black")+
  geom_polygon(data = usa,
               aes(x = long,
               y = lat,
               group = group),
               fill = NA,
               color = "black")+
  theme_map()+
  coord_fixed(1.3)

akc_hawaii <- ggplot()+
  geom_polygon(data = hawaii,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = NA,
               color = "black")+
  geom_polygon(data = hawaii_oahu,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = "black")+
  geom_polygon(data = other_islands,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = "gray")+
  theme_map()+
  coord_fixed(1.3)
akc_usa + akc_hawaii + plot_layout(ncol = 2, widths = c(1.5, 1))
```
#Hmm.  The big island of Hawaii didn't show up as gray.  I guess the name in map_data is different.  The code is there that would select it.  On to Part 2!

***
###Part 2: Dot distribution maps
***
###Load the libraries 
#already loaded
***
###Define some variables. set.seed value as in assignment so I can see if I'm doing it right! 
```{r}
min_long <- -126
max_long <- -104
min_lat <- 31
max_lat <- 50

set.seed(9618973)
file_path <-"~/Documents/BI685Files/GitHubDocs/chronister_andrew/data"
```

***
###Read in the "Gigantopithecus" data
```{r}
bigfoot <- read.csv(file.path(file_path, "bigfoot.csv"))

bigfoot <- filter(bigfoot,
                  long >= min_long & long <= max_long,
                  lat >= min_lat & lat <= max_lat) %>%
  sample_n(300) %>%
  mutate(name = "Gigantiopithecus") %>%
  select(name, long, lat)
```
#Read in the bears data
```{r}
bears <- read_csv(file.path(file_path, "bears.csv"))

both_species <- bind_rows(bigfoot, bears)

head(both_species)
```
```{r}
tail(both_species)
```
#Get terrain map
```{r}
base = get_stamenmap(bbox = c(min_long,
                              min_lat,
                              max_long,
                              max_lat),
                     zoom = 4,
                     maptype = "terrain-background")
ggmap(base)
```
```{r}
base <- ggmap(base)

base + geom_point(data = bigfoot,
                  aes(x = long,
                      y = lat))
```
```{r}
base + geom_point(data = bears,
                  aes(x = long,
                  y = lat))
```
#Both species
```{r}
base + geom_point(data = both_species,
                  aes(x = long,
                      y = lat,
                      shape = name,
                      fill = name),
                  size = 2)+
  scale_shape_manual(values = c(22:23))+
  labs(x = "Longitude",
       y = "Latitude",
       fill = "Species",
       shape = "Species")
```
***
###Recreate with SO number as seed.  
```{r}
set.seed(1065450)
#I could just redo everything from here with the existing code and should get the new results.  
bigfoot <- read_csv(file.path(file_path, "bigfoot.csv"))
bigfoot <- filter(bigfoot,
                   long >= min_long & long <= max_long,
                   lat >= min_lat & lat <= max_lat) %>%
  sample_n(300) %>%
  mutate(name = "Gigantopithecus") %>%
  select(name, long, lat)
```
#Bears data already read in.  
```{r}
both_species <- bind_rows(bigfoot, bears)

head(both_species)
tail(both_species)
```
```{r}
base = get_stamenmap(bbox = c(min_long,
                              min_lat,
                              max_long,
                              max_lat),
                     zoom = 4,
                     maptype = "terrain-background")

base <- ggmap(base)
base + geom_point(data = both_species,
                  aes(x = long,
                      y = lat,
                      shape = name,
                      fill = name),
                  size = 2)+
  scale_shape_manual(values = c(22:23))+
  labs(x = "Longitude",
       y = "Latitude",
       fill = "Species",
       shape = "Species")
```
***
###Part3: Bathymetry
***
###Load the libraries
#Already loaded
***
###Define the variables
```{r}
min_long <- -170
max_long <- 164
min_lat <- -42
max_lat <- -8

earthquakes <- quakes %>%
  mutate(depth = depth * 3280.84)
```
#load NOAA data
```{r}
fiji <- getNOAA.bathy(lon1 = min_long,
                      lon2 = max_long,
                      lat1 = min_lat,
                      lat2 = max_lat,
                      antimeridian = TRUE,
                      keep = TRUE)
```

```{r}
autoplot.bathy(fiji,
               geom = c("raster", "contour"),
               size = 0.3,
               na.rm = TRUE)
```
```{r}
autoplot(fiji,
         geom = c("raster", "contour"),
         size = 0.1,
         na.rm = TRUE)+
  scale_fill_gradient2(low = "dodgerblue4",
                       mid = "gainsboro",
                       high = "darkgreen",
                       name = "Depth (ft)")+
  labs(x = "Longitude",
       y = "Latitude")+
  theme(axis.title.y = element_text(angle = 0,
                                    vjust = 0.5))
```
```{r}
base_map <- autoplot(fiji,
         geom = c("raster", "contour"),
         size = 0.1,
         na.rm = TRUE)+
  scale_fill_gradient2(low = "dodgerblue4",
                       mid = "gainsboro",
                       high = "darkgreen",
                       name = "Depth (ft)")+
  labs(x = "Longitude",
       y = "Latitude")+
  theme(axis.title.y = element_text(angle = 0,
                                    vjust = 0.5))
```
```{r}
base_map + geom_point(data = earthquakes,
                      aes(x = long,
                      y = lat))
```
#Use alpha arguement
```{r}
base_map + geom_point(data = earthquakes,
                      aes(x = long,
                          y = lat,
                          size = mag),
                      alpha = 0.4)
```
#scale size a little smaller
```{r}
base_map + geom_point(data = earthquakes,
                      aes(x = long,
                          y = lat,
                          size = mag),
                      alpha = 0.4) +
  scale_size_continuous(range = c(0.5, 3), name = "Magnitude")
```
***
###Now you try it.
***
###Set new variables
```{r}
min_long <- -90
max_long <- -58
min_lat <- 8
max_lat <- 28
```
#Read in the data
```{r}
blennies <- read_csv("data/blennies.csv")
blennies
```
#Get the caribbean data
```{r}
carib_sea <- getNOAA.bathy(lon1 = min_long,
                           lon2 = max_long,
                           lat1 = min_lat,
                           lat2 = max_lat,
                           antimeridian = FALSE,
                           keep = TRUE)
```
#Use autoplot for the base map
```{r}
base_map <- autoplot(carib_sea,
         geom = c("raster", "contour"),
         size = 0.1,
         na.rm = TRUE) +
  scale_fill_etopo(guide = FALSE)+
  labs(x = "Longitude",
       y = "Latitude")+
  theme(axis.title.y = element_text(angle = 0,
                                    vjust = 0.5))
base_map
```
#Add the blennies
```{r}
base_map + geom_point(data = blennies,
                      aes(x = long,
                          y = lat,
                          shape = name),
                      fill = "white",
                      size = 3.2)+
  scale_shape_manual(values = c(21:23))+
  labs(shape = "Species")+
  theme(legend.position = "bottom")+
  theme(legend.text = element_text(face = "italic"))

```



