---
title: "HW 06 Part 1: Complete the sections"
author: "Andrew Chronister"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- Change "your name" in the YAML header above to your name.

- As usual, enter the examples in code chunks and run them, unless told otherwise.

## Chapter 5: Data transformation

Read [R4ds Chapter 5: Data transformation](https://r4ds.had.co.nz/transform.html), sections 5-7.
#setup
```{r}
library(tidyverse)
library(nycflights13)

```

You do not answer any of the questions this week. Just read and run all of the examples.

### Section 5.5: Add new variables with `mutate()`

Type and run all of the examples.
```{r}
flights_sml <- select(flights,
                      year:day,
                      ends_with("delay"),
                      distance,
                      air_time
                      )
mutate(flights_sml,
       gain = dep_delay - arr_delay,
       speed = distance / air_time * 60
       )
```
#This could have come in useful with a question in the last assignment. I was struggling with how to do just this!  :D
```{r}
mutate(flights_sml,
       gain = dep_delay - arr_delay,
       hours = air_time / 60,
       gain_per_hour = gain / hours
       )
```

```{r}
transmute(flights,
          gain = dep_delay - arr_delay,
          hours = air_time / 60,
          gain_per_hour = gain / hours
          )
```

Pay careful attention to the useful creation functions in section 5.5.1. 
```{r}
transmute(flights,
          dep_time,
          hour = dep_time %/% 100,
          minute = dep_time %% 100
          )
```
```{r}
(x <- 1:10)
lag(x)
lead(x)
```
```{r}
x
cumsum(x)
cummean(x)
```
```{r}
y <- c(1,2,2,NA,3,4)
min_rank(y)
min_rank(desc(y))
```

```{r}
row_number(y)
dense_rank(y)
percent_rank(y)
cume_dist(y)
```

#Skipping the 5.5.2 questions as instructed.

### Section 5.6: Grouped summaries with `summarise()`

Read and run all of the examples. Pay careful attention to the useful summary functions. Unless you really like baseball, skip the baseball part of 5.6.3.
```{r}
summarize(flights, delay = mean(dep_delay, na.rm = TRUE))
```

```{r}
by_day <-  group_by(flights, year, month, day)
summarize(by_day, delay = mean(dep_delay, na.rm = TRUE))
```

#5.6.1
```{r}
by_dest <- group_by(flights, dest)
delay <- summarize(by_dest,
                   count = n(),
                   dist = mean(distance, na.rm = TRUE),
                   delay = mean(arr_delay, na.rm = TRUE)
                   )
delay <- filter(delay, count>20, dest != "HNL")
ggplot(data = delay, mapping = aes(x=dist, y=delay))+
  geom_point(aes(size = count), alpha = 1/3)+
  geom_smooth(se = FALSE)
```
#Pipefest time...
```{r}
delays <- flights %>%
  group_by(dest) %>%
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>%
  filter(count>20, dest != "HNL")
```

### Section 5.7: Grouped mutates (and filters)

Read and run all of the examples.

