---
title: "HW 06 Part 1: Complete the sections"
author: "Andrew Chronister"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- Change "your name" in the YAML header above to your name.

- As usual, enter the examples in code chunks and run them, unless told otherwise.

## Chapter 5: Data transformation

Read [R4ds Chapter 5: Data transformation](https://r4ds.had.co.nz/transform.html), sections 5-7.
#setup
```{r}
library(tidyverse)
library(nycflights13)

```

You do not answer any of the questions this week. Just read and run all of the examples.

### Section 5.5: Add new variables with `mutate()`

Type and run all of the examples.
```{r}
flights_sml <- select(flights,
                      year:day,
                      ends_with("delay"),
                      distance,
                      air_time
                      )
mutate(flights_sml,
       gain = dep_delay - arr_delay,
       speed = distance / air_time * 60
       )
```
#This could have come in useful with a question in the last assignment. I was struggling with how to do just this!  :D
```{r}
mutate(flights_sml,
       gain = dep_delay - arr_delay,
       hours = air_time / 60,
       gain_per_hour = gain / hours
       )
```

```{r}
transmute(flights,
          gain = dep_delay - arr_delay,
          hours = air_time / 60,
          gain_per_hour = gain / hours
          )
```

Pay careful attention to the useful creation functions in section 5.5.1. 
```{r}
transmute(flights,
          dep_time,
          hour = dep_time %/% 100,
          minute = dep_time %% 100
          )
```
```{r}
(x <- 1:10)
lag(x)
lead(x)
```
```{r}
x
cumsum(x)
cummean(x)
```
```{r}
y <- c(1,2,2,NA,3,4)
min_rank(y)
min_rank(desc(y))
```

```{r}
row_number(y)
dense_rank(y)
percent_rank(y)
cume_dist(y)
```

#Skipping the 5.5.2 questions as instructed.

### Section 5.6: Grouped summaries with `summarise()`

Read and run all of the examples. Pay careful attention to the useful summary functions. Unless you really like baseball, skip the baseball part of 5.6.3.
```{r}
summarize(flights, delay = mean(dep_delay, na.rm = TRUE))
```

```{r}
by_day <-  group_by(flights, year, month, day)
summarize(by_day, delay = mean(dep_delay, na.rm = TRUE))
```

#5.6.1
```{r}
by_dest <- group_by(flights, dest)
delay <- summarize(by_dest,
                   count = n(),
                   dist = mean(distance, na.rm = TRUE),
                   delay = mean(arr_delay, na.rm = TRUE)
                   )
delay <- filter(delay, count>20, dest != "HNL")
ggplot(data = delay, mapping = aes(x=dist, y=delay))+
  geom_point(aes(size = count), alpha = 1/3)+
  geom_smooth(se = FALSE)
```
#Pipefest time...
```{r}
delays <- flights %>%
  group_by(dest) %>%
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>%
  filter(count>20, dest != "HNL")
```

```{r}
flights %>%
  group_by(year, month, day) %>%
  summarise(mean = mean(dep_delay))
```

```{r}
flights %>%
  group_by(year, month, day) %>%
  summarize(mean = mean(dep_delay, na.rm = TRUE))

```

```{r}
not_cancelled <-flights %>%
  filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled %>%
  group_by(year, month, day) %>%
  summarise(mean = mean(dep_delay))
```
#Section 5.6.3
```{r}
delays <- not_cancelled %>%
  group_by(tailnum) %>%
  summarise(
    delay = mean(arr_delay)
  )
ggplot(data=delays, mapping = aes(x = delay))+
  geom_freqpoly(binwidth = 10)
```

```{r}
delays <-  not_cancelled %>%
  group_by(tailnum) %>%
  summarise(
    delay=mean(arr_delay, na.rm = TRUE),
    n=n()
  )
ggplot(data = delays, mapping = aes(x=n, y=delay))+
  geom_point(alpha = 1/10)

```

```{r}
delays %>%
  filter(n>25) %>%
  ggplot(mapping = aes(x=n, y=delay))+
  geom_point(alpha = 1/10)
```
#Skipping the baseball reference.

#Section 5.6.4
```{r}
not_cancelled %>%
  group_by(year, month, day) %>%
  summarize(
    avg_delay = mean(arr_delay),
    avg_delay2 = mean(arr_delay[arr_delay>0])
  )
```
```{r}
not_cancelled %>%
  group_by(dest) %>%
  summarize(distance_sd = sd(distance)) %>%
  arrange(desc(distance_sd))
```

```{r}
not_cancelled %>%
  group_by(year, month, day)%>%
  summarise(
    first = min(dep_time),
    last = max(dep_time)
  )
```
```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    first_dep = first(dep_time), 
    last_dep = last(dep_time)
  )
```
```{r}
not_cancelled%>%
  group_by(year, month, day)%>%
  mutate(r = min_rank(desc(dep_time))) %>%
  filter(r %in% range(r))
```
```{r}
not_cancelled %>%
  group_by(dest) %>%
  summarise(carriers = n_distinct(carrier)) %>%
  arrange(desc(carriers))
```

```{r}
not_cancelled %>%
  count(dest)
```
```{r}
not_cancelled%>%
  count(tailnum, wt = distance)
```

```{r}
not_cancelled%>%
  group_by(year, month, day)%>%
  summarise(n_early = sum(dep_time<500))
not_cancelled%>%
  group_by(year, month, day)%>%
  summarise(hour_perc = mean(arr_delay > 60))
```

```{r}
daily <- group_by(flights, year, month, day)
(per_day <- summarise(daily, flights = n()))
(per_month <- summarise(per_day, flights = sum(flights)))
(per_year <- summarise(per_month, flights = sum(flights)))
```
```{r}
daily %>%
  ungroup() %>%
  summarize(flights = n())
```

#skipping questions 5.6.7 as instructed.

### Section 5.7: Grouped mutates (and filters)
```{r}
flights_sml %>% 
  group_by(year, month, day) %>%
  filter(rank(desc(arr_delay)) < 10)
```

Read and run all of the examples.
```{r}
popular_dests <- flights %>% 
  group_by(dest) %>% 
  filter(n() > 365)
popular_dests
```
```{r}
popular_dests %>% 
  filter(arr_delay > 0) %>% 
  mutate(prop_delay = arr_delay / sum(arr_delay)) %>% 
  select(year:day, dest, arr_delay, prop_delay)
```
#Skipping exercises 5.7.1
