---
title: "HW 06 Part 2: Wrangling II"
author: "Andrew Chronister"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
***
###Load tidyverse and knitr libraries
```{r}
library(tidyverse)
library(knitr)
install.packages("smatr")
library(smatr)
```
***
###Defining file_path variable
```{r}
file_path <- "~/Documents/BI685Files/GitHubDocs/chronister_andrew/data/"
```
***
###Read in darters.txt (In reviewing I noticed that I DIDN'T include errant spaces in the column names this time.  Hooray for me!)
```{r}
start_pos <-  c(1,4,15,16,17,18,21,25,29,33,36,39,41)
stop_pos <- c(3,14,15,16,17,20,24,28,32,35,38,40,42)
names <- c("specimen id", "species", "riffle", "age", "sex", "length", "depth", "velocity", "temperature", "majsub", "minsub", "majtype", "mintype")
darters <- read_fwf(file.path(file_path, "darter.txt"), fwf_positions(start_pos, stop_pos, col_names = names), skip = 13)
darters
```
***
###remove tetrazonum and zonale
```{r}
darters2 <- darters %>%
  filter(!species %in% c("tetrazonum", "zonale"))
darters2
```
***
###Wrangle the data (First time I did this, for some reason, I thought we were supposed to wrangle each step separately, not accumulate the changes as we went on.  My mistake!)
```{r}
  mutate(darters2,
       total_substrate = majsub + minsub
       )
```
#My calculations show that total_substrate = 1 for all observations.  

```{r}
darters2 <- mutate(darters2,
       total_substrate = majsub + minsub
       )
darters2
```
#Change riffle values
```{r}
darters2 <- mutate(darters2,
         riffle = ifelse(riffle == 1,
                         "Riffle 1",
                         "Riffle 2"))
darters2
```
#Change length data from cm to mm
```{r}
darters2 <- mutate(darters2,
         length = length * 10)
darters2
```
#Change sex values
```{r}
darters2 <- mutate(darters2,
         sex = ifelse(sex == "f",
                      "Female",
                      "Male"))
darters2
```
#Summarize data
```{r}
darters_species_means <- darters2 %>%
  group_by(species, riffle) %>%
  summarize(length = mean(length, na.rm = TRUE),
            depth = mean(depth, na.rm = TRUE),
            velocity = mean(velocity, na.rm = TRUE)
            )
darters_species_means
```
***
###Plot 1
#make a plot
```{r}
ggplot(data = darters2) +
  stat_summary(mapping = aes(x=species, y=length),
               fun.ymin = min,
               fun.ymax = max,
               fun.y = median
               ) +
  geom_point(darters_species_means, mapping = aes(x=species, y=length), shape = 15)
  
```
#Turn plot
```{r}
ggplot(data = darters2) +
  stat_summary(mapping = aes(x=species, y=length),
               fun.ymin = min,
               fun.ymax = max,
               fun.y = median
               ) +
  geom_point(data = darters_species_means, mapping = aes(x=species, y=length), shape = 15)+
  coord_flip() 
  
```
#Add Facet wrap
```{r}
ggplot(data = darters2) +
  stat_summary(mapping = aes(x=species, y=length),
               fun.ymin = min,
               fun.ymax = max,
               fun.y = median
               ) +
  geom_point(data = darters_species_means, mapping = aes(x=species, y=length), shape = 15)+
  coord_flip()+
  facet_wrap(~ riffle, nrow = 2)
```
#Change labels
```{r}
ggplot(data = darters2) +
  stat_summary(mapping = aes(x=species, y=length),
               fun.ymin = min,
               fun.ymax = max,
               fun.y = median
               ) +
  geom_point(data = darters_species_means, mapping = aes(x=species, y=length), shape = 15)+
  coord_flip()+
  facet_wrap(~ riffle, nrow = 2)+
  labs(x = NULL, y = "Length (cm)")
```


#The mean size of of the spectabile and caeruleum species are significantly larger in riffle 1 than riffle 2. the blennioides species is the opposite, with modestly larger size in riffle 1 vs. riffle 2.

***
###Plot2
#initial plot
```{r}
ggplot(data = darters2, mapping = aes(x=length, y=species))+
  geom_point(color = "grey70")
```
#add stat_summary
```{r}
ggplot(data = darters2, mapping = aes(x=length, y=species))+
  geom_point(color="grey70")+
  stat_summary(data = darters_species_means, mapping = aes(x=length, y=species),
               fun.y = median,
               geom = "point",
               size = 2.2
               )
```
#add second stat_summary
```{r}
ggplot(data = darters2, mapping = aes(x=length, y=species))+
  geom_point(color="grey70")+
  stat_summary(data = darters_species_means, mapping = aes(x=length, y=species),
               fun.y = median,
               geom = "point",
               size=2.2
               )+
  stat_summary(data = darters_species_means, mapping = aes(x=length, y=species),
               fun.y = mean,
               geom = "point",
               size=3,
               color = "maroon"
               )
```
#flip, facet, label
```{r}
ggplot(data = darters2, mapping = aes(x=length, y=species))+
  geom_point(color="grey70")+
  stat_summary(data = darters_species_means, mapping = aes(x=length, y=species),
               fun.y = median,
               geom = "point",
               size=2.2
               )+
  stat_summary(data = darters_species_means, mapping = aes(x=length, y=species),
               fun.y = mean,
               geom = "point",
               size=3,
               color = "maroon"
               )+
  coord_flip()+
  facet_wrap(~ riffle, nrow=2)+
  labs(y=NULL, x="Length (cm)")
```
***
###Stress and corticosterones in birds

***
###Load the libraries
```{r}
library(tidyverse)
library(smatr)
```
***
###Define the filepath and read in the data
```{r}
file_path <- "~/Documents/BI685Files/GitHubDocs/chronister_andrew/data/"
names <- c("capture_date", "day", "site", "urban_score", "cortF", "cortB", "sex", "smi", "tarsus", "mass", "basal_cort", "stress_cort")
feathers <- read_csv(file.path(file_path, "feathers.csv"), locale = locale(decimal_mark = ","), skip = 1, col_names = names)
feathers
```
#the data is tidy. Each column is a variable, each row is an observation, and each cell is a value.

#convert decimal mark, change column names done above

***
###Wrangle the data
#calculate mean tarsus
```{r}
mean_tarsus <- mean(feathers[["tarsus"]])
mean_tarsus
```
#use mutate to transform
```{r}
feathers2 <- mutate(feathers,
       log_mass = log(mass),
       log_tarsus = log(tarsus)
       )
feathers2

```
#Run function
```{r}
major_axis <- sma(log_mass ~ log_tarsus, data = feathers2)
major_axis
```
#Note, I have no idea what those numbers mean.  Wow... just wow.  :D

#Apply the summary function
```{r}
summary(major_axis)
```
#Apply the coef function
```{r}
b <- coef(major_axis)["slope"]
b
```
#put it all together to figure out SMI
```{r}
feathers3 <- mutate(feathers2,
       new_smi = mass * (mean_tarsus / tarsus)^b)

select(feathers3, smi, new_smi)
```
#smi and new_smi are very similar. now on to summarize...

***
###Summarize Data
```{r}
site_data <- feathers2 %>%
  group_by(site)%>%
  summarize(basal_cort_mean = mean(basal_cort, na.rm = TRUE),
            basal_cort_sde = sd(basal_cort, na.rm = TRUE)/sqrt(43),
            stress_cort_mean = mean(stress_cort, na.rm = TRUE),
            stress_cort_sde = sd(stress_cort, na.rm = TRUE)/sqrt(43) 
  )
site_data
```
#I'm afraid that the values for the standard deviation error are incorrect because of the NA values.  Gonna run this again, a little differently.

```{r}
sum(is.na(feathers2$basal_cort))
sum(is.na(feathers2$stress_cort))
```
#Unless I'm overthinking it, the values for the sqrt should be "33" and "40" accordingly. That's what I'm gonna do... even if it is wrong. :D Or I could have removed the NA values when tidying the data... that would have been smart. FIXED AS PER SUGGESTIONS! 

```{r}
site_data <- feathers2 %>%
  group_by(site) %>%
  summarize(n=n(), 
            basal_cort_mean = mean(basal_cort, na.rm = TRUE),
            basal_cort_sde = sd(basal_cort, na.rm =TRUE)/sqrt(n),
            stress_cort_mean = mean(stress_cort, na.rm = TRUE),
            stress_cort_sde =sd(stress_cort, na.rm = TRUE)/sqrt(n)
            )
site_data
```
***
###Graph the Data
#Make a boxplot of new_smi
```{r}
  ggplot(data = feathers3, mapping = aes(x=site, y=new_smi)) +
  geom_boxplot()
```
#the scaled mass index for site NIORT has a huge variation between the max and min as shown by the whiskers of the boxplot.  Also, the villefollet site has lower values than the rest of the locations.

##Plot 2 (had to redefine basalcortmean and basalcortsde because I didn't save them to variables earlier.  I guess I could do this now, pretty easily, in reviewing the assignment... but I'm gonna let it stay as an example of how to do way more coding than necessary.) FIXING IT AS PER SUGGESTION. much better! Much better TIE fighters! 
```{r}
ggplot(data=site_data,
       mapping = aes(x=site, y=basal_cort_mean))+
  geom_point()+
  geom_errorbar(aes(ymin = basal_cort_mean - basal_cort_sde,
                    ymax = basal_cort_mean + basal_cort_sde))
```
#Plot 3-repeat the plot without tie fighters

```{r}
ggplot(data = site_data,
       mapping = aes(x=site, y=stress_cort_mean))+
  geom_point()+
  geom_linerange(aes(ymin = stress_cort_mean - stress_cort_sde,
                     ymax = stress_cort_mean + stress_cort_sde))
```
#I'm happy about this now!   I knew something was terribly, horribly wrong. :D

#<old text>I'm not happy about one thing.  The errorbar or linerange data is the same for each site.  I am thinking that it should be different at each site.  I'm still thinking about how that could be done, or if that is what I should be trying to do...

####Grad Student Extras...
#use size 3 point located at the mean
```{r}
ggplot(data = site_data, mapping = aes(x=site, y=stress_cort_mean))+
  geom_point()+
  geom_point(mapping = aes(x=site, y=stress_cort_mean), size = 3)+
  geom_linerange(aes(ymin = stress_cort_mean - stress_cort_sde,
                     ymax = stress_cort_mean + stress_cort_sde))
```

#Make it maroon
```{r}
ggplot(data = site_data, mapping = aes(x=site, y=stress_cort_mean))+
  geom_point()+
  geom_point(mapping = aes(x=site, y=stress_cort_mean), size = 3, color = "maroon")+
  geom_linerange(aes(ymin = stress_cort_mean - stress_cort_sde,
                     ymax = stress_cort_mean + stress_cort_sde))
```
#Place error bars behind points, also changed size and alpha of linerange for viewability

```{r}
ggplot(data = site_data, mapping = aes(x=site, y=stress_cort_mean))+
  geom_linerange(aes(ymin = stress_cort_mean - stress_cort_sde,
                     ymax = stress_cort_mean + stress_cort_sde), alpha = 1/5,
                     size = 2)+
  geom_point()+
  geom_point(mapping = aes(x=site, y=stress_cort_mean), size = 3, color = "maroon")
```
#Make the TIE fighters look more like TIE fighters!
```{r}
ggplot(data = site_data, mapping = aes(x=site, y=stress_cort_mean))+
  geom_errorbar(aes(ymin = stress_cort_mean - stress_cort_sde,
                     ymax = stress_cort_mean + stress_cort_sde), alpha = 1/5,
                     size = 1/2)+
  geom_point()+
  geom_point(mapping = aes(x=site, y=stress_cort_mean), size = 3, color = "maroon")
```
#Now THOSE are some beautiful TIE fighters!  




